Comparing Apples and Oranges in Youth Depression Treatments?

### 2024-08-08 multilevel_metaregression/manuscript_with_multilevel_metan.qmd
added sensitivity analyses and sex meta-analysis with all studies in supplement

# For referencing
bib_data <- ReadBib("emotion_concepts_paper.bib")

@@

df_appl_v_orange <-readxl:: read_excel("df_appl_v_orange.xlsx")

@@

df_scale_ranges <- read.csv("scale_ranges.csv")
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

@@

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample
# E) the CDRS only sample (**Argyris** do we still want this?)
# F) Clinical Levels Depression

@@

# D Clinical Levels Depression Sensitivity Analysis ---------------------------------------

@@

# E. HAMD Sensitivity Analysis ---------------------------------------

@@

@@

Data for included are summarised in @tbl-all-trials and are also available as a csv dataframe on \[<https://github.com/transatlantic-comppsych/apples_oranges>\]. Please see @fig-prisma for a summary of the sources of included RCTs.

@@

First, the grounds for comparison between medication and psychotherapy should be seen as shaky, rather than offering confidence, and there is an urgent need to revisit guidelines and public information in light of the limitations.

>>>>

# For referencing
bib_data <- ReadBib("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/emotion_concepts_paper.bib")

@@

df_appl_v_orange <-readxl::read_excel("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/df_appl_v_orange.xlsx")

@@

df_scale_ranges <- read.csv("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/scale_ranges.csv")
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

@@

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample
# D) clinical levels of depression
# E) the CDRS only sample
# E) HAMD sensitivity analysis
# G) Studies with post SD
# H) Studies with variance <.02

@@

# D. Clinical Levels Depression Sensitivity Analysis---------------------------------------

@@

# F. HAMD Sensitivity Analysis ---------------------------------------

@@

# G. Post SD Sensitivity Analysis ---------------------------------------

##--------- Data preparation

#For this we need to filter the master df by missing post_sd_active and post_sd_control in df_first_row

#create subset of first_row

df_first_row <- read.csv("C:/Users/Raphaelle/Documents/AIM Lab/apples_and_oranges/main_git/apples_oranges/df_first_row.csv")

#create new_study_ids to match master dataset
df_first_row  <- df_first_row  %>%
  mutate(new_study_id = case_when(psy_or_med == 0 ~ paste(study,year, sep = ", "),
                                  .default = study ))

df_first_row_reduced <- df_first_row %>%
  filter(!(new_study_id == "March, 2004" & is.na(control_type))) %>%
  select(new_study_id, post_sd_active, post_sd_control)

#as first_row has multiple rows per study (showing different arm pairs), I will keep only one row per study

df_first_row_reduced <- df_first_row_reduced %>%
  group_by(new_study_id) %>%
  slice_head(n = 1) %>%
  ungroup()

#create a column in master df that shows 1 if any missing post sds in first_row

df_long_for_metan_with_missing <- df_long_for_metan %>%
  left_join(df_first_row_reduced, by = "new_study_id") %>%
  mutate(missing_post_sd = ifelse(is.na(post_sd_active) | is.na(post_sd_control), 1, 0)) %>%
  select(-post_sd_active, -post_sd_control) %>%
  arrange(arm_effect_size, new_study_id)

#all studies missing sds in active are also missing them in control, so we can go ahead and filter out all arms of these studies

df_long_for_metan_without_missing <- df_long_for_metan_with_missing %>%
  filter(missing_post_sd == 0)

n_distinct(df_long_for_metan_without_missing$new_study_id) #61 studies
n_distinct(df_long_for_metan$new_study_id) #88 studies total
nrow(df_long_for_metan_without_missing) #140 arms
nrow(df_long_for_metan) #203 arms

##------------ Estimates

#take study ids where study does not have missing post SD
sd_study_ids <-
  df_long_for_metan_with_missing[df_long_for_metan_with_missing$missing_post_sd ==0,]$new_study_id

# use those ids to subset all dataframes in the list
list_sd_study <- list()

  list_sd_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                 %in% sd_study_ids,]
}

# count the number of studies
study_count_sd_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% sd_study_ids,])

# apply the metareg function to the list of clinical studies
model_1 <- as.formula(~ four_level_var)
list_model_1_sd_study <- run_multi_metaregression(list_sd_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_sd_study <-  aggregate_model_results(list_model_1_sd_study, condition)

# extract SMDs from the list for clinical studies
means_sd_study <-  lapply(list_model_1_sd_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_sd_study <- calculate_mean_coefs_ses(means_sd_study, condition)
coef_and_se_means_sd_study <- data.frame(cbind(condition, coef_and_se_means_sd_study))

# add the ns to the dataframe
coef_and_se_means_sd_study <- left_join(coef_and_se_means_sd_study, study_count_sd_study[study_count_sd_study$is_missing == FALSE, ] %>% 
                                            ungroup() %>% 
                                            select(condition, n),  
                                          by = "condition")

#H. Studies with variances <.02-----------------------------

##-----------Data preparation

#create column with variance in master df
df_long_for_metan$baseline_variances <- (df_long_for_metan$perc_baseline_sd)^2

#take study ids where variance =2 or less
var2_study_ids <-
  df_long_for_metan[df_long_for_metan$baseline_variances<=0.02,]$new_study_id
length(var2_study_ids)

##------------Estimates

# use those ids to subset all dataframes in the list
list_var2_study <- list()

  list_var2_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                               %in% var2_study_ids,]
}

# count the number of studies
study_count_var2_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% var2_study_ids,])

# apply the metareg function to the list of clinical studies
model_1 <- as.formula(~ four_level_var)
list_model_1_var2_study <- run_multi_metaregression(list_var2_study, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_var2_study <-  aggregate_model_results(list_model_1_var2_study, condition)

# extract SMDs from the list for clinical studies
means_var2_study <-  lapply(list_model_1_var2_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_var2_study <- calculate_mean_coefs_ses(means_var2_study, condition)
coef_and_se_means_var2_study <- data.frame(cbind(condition, coef_and_se_means_var2_study))

# add the ns to the dataframe
coef_and_se_means_var2_study <- left_join(coef_and_se_means_var2_study, study_count_var2_study[study_count_var2_study$is_missing == FALSE, ] %>% 
                                          ungroup() %>% 
                                          select(condition, n),  
                                        by = "condition")

@@

Data for included studies are summarised in @tbl-all-trials and are also available as a csv dataframe on \[<https://github.com/transatlantic-comppsych/apples_oranges>\]. Please see @fig-prisma for a summary of the sources of included RCTs.

@@

First, the grounds for comparison between medication and psychotherapy should be seen as shaky, rather than offering confidence, and there is a need to revisit guidelines and public information in light of the limitations.
